### 2024-08-05 manuscript.tex
Various changes and additions

\begin{center} \Large \sffamily\textbf{Using ship observations to assess Southern Ocean clouds in a storm-resolving general circulation model ICON}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Currently, a new generation of km-scale (also called `k-scale') resolution
global climate models are in development as the forthcoming phase of climate
modelling. One such model is a 5-km version of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON) developed jointly by Deutscher Wetterdienst
(DWD) and the Max-Planck-Institute for Meteorology (MPI-M). Because of the high
resolution, most parametrisations, such as that of convection and clouds, can
be avoided.

Previous studies have identified substantial large-scale biases in model clouds
over the Southern Ocean (SO), affecting sea surface temperature and the Earth's
albedo overall. Our aim is to quantify how well the high-resolution ICON model
is simulating clouds in this region, particularly in light of the fact that
subgrid-scale clouds are not parametrised in this model. This region is mostly
dominated by boundary layer clouds generated by shallow convection, and these
are problematic to observe by spaceborne lidar and radars, which are affected
by attenuation by overlapping and thick clouds and ground clutter,
respectively. Therefore, we choose to use a large set of ship-based
observations conducted with ceilometers and lidars on board of Research Vessel
(RV) \textit{Polarstern} and other voyages. Altogether, we analyse over 2000
days (about 6 years) of data from 31 voyages and 1 sub-antarctic station
covering diverse longitudes of the SO. To achieve a like-for-like comparison
with the model, we use a ground-based lidar simulator called the Automatic
Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF
Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and
Applications, Version 2 (MERRA-2).

We show that the model underestimates the total cloud fraction by about 10\%,
with overestimation of cloud below 2 km, and underestimation of cloud above 2
km. The reanalyses also underestimate the total cloud fraction by about 20\%.
ERA5 overestimates cloud below 1 km but underestimates near-surface cloud or
fog. In addition to lidar data, we compare radiosonde profiles acquired on the
RV \textit{Polarstern} voyages with ICON. Notably, the model exhibits smaller
natural variability than observations, and its lifting condensation level tends
to be higher. This might explain why cloud occurrence is peaking higher in the
model (at 500 m) than in observations (at the surface).

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/map-fig.png}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern}, RSV
\emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer}, and
HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010 to 2021.
The dashed line at 40$^\circ$S delineates the Southern Ocean area of our
analysis. A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke
Mehrtens/Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma), \textbf{(d)} Vaisala CL51 install at the
Macquarie Island station (© Jeﬀ Aquilina, Bureau of Meteorology).
% TODO: Ask for photo permission and info about license.
}
\label{fig:map}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/lts-dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data. Shown is also the chosen
dividing threshold of 12 K for `stable' and `unstable' conditions.
}
\label{fig:example}
\end{figure}

@@

These data sources contained ceilometer (lidar) observations with the Vaisala
CL51 operating at 910 nm and the Lufft CHM-15k operating at 1064 nm. A
ceilometer is a low-power vertically pointing lidar with the purpose of
measuring the cloud base, but it also measures the full vertical structure of
clouds as long as the laser signal is not attenuated by thick clouds. A total
of about 1500 days of lidar observations were included in our analysis.

@@

\subsection{Vaisala CL51}

The Vaisala CL51 is a ceilometer operating at a near-infrared wavelength of 910
nm. The maximum range 15.4 km, vertical resolution 10 m, and the number of
vertical levels is 1540. The wavelength is affected by water vapour absorption
of about 20\% in the mid-latitudes \citep{wiegner2015,wiegner2019}.  The
instrument data files containing uncalibrated backscatter were first converted
to NetCDF with cl2nc (\url{https://github.com/peterkuma/cl2nc}) and then
processed with the ALCF to produce a cloud mask and cloud occurrence by height
and the total cloud fraction.

@@

@@

ERA5 \citep{era5} is a reanalysis produced by the European Centre for
Medium-Range Weather Forecasts (ECMWF). It is based on a numerical weather
prediction (NWP) model Integrated Forecast System (IFS) version CY41R2.  The
horizontal resolution is 0.25$^\circ$ in latitude and longitude (about 28 km in
the North--South direction and 14 km in the East--West direction at
60$^\circ$S). The model uses 137 vertical levels.

@@

\begin{table}
% TODO: Add sampling intervals.
\caption{List of voyages and stations. Start, end and days (UTC; inclusive) indicate the time period when the vessel was south of 40$^\circ$S. Abbreviations: ceilometer (ceil.), Australia (AU), New Zealand (NZ), South America (SA), Atlantic Ocean (AO) and Africa (AF). The number of days is rounded to the nearest integer.}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
Name & Vessel & Ceil. & Region & Start & End & Days\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124\\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CL51    & AU       & 2017-10-29 & 2018-03-26 & 149\\
MCQ16-18 & /                             & CL51    & AU/NZ    & 2016-04-03 & 2018-03-14 & 710\\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51    & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &          &            &            &         & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}

>>>>

\begin{center} \Large \sffamily\textbf{Using ship observations to assess Southern Ocean clouds in the storm-resolving general circulation model ICON}\\[0.4cm]
\large Peter Kuma$^\mathsf{1,}$*, Frida A.-M. Bender$^\mathsf{1}$, Adrian J. McDonald$^\mathsf{2}$ and Thorsten Mauritsen$^\mathsf{1}$\\[0.4cm]
\small
$^\mathsf{1}$Department of Meteorology (MISU) and Bolin Centre for Climate Research, Stockholm University, Stockholm SE-106 91, Sweden\\[0.2cm]
$^\mathsf{2}$School of Physical and Chemical Sciences, University of Canterbury, Christchurch, New Zealand\\[0.2cm]
*Corresponding author: Peter Kuma (\href{mailto:peter.kuma@misu.su.se}{peter.kuma@misu.su.se})\\[0.4cm]
\large \today\\[0.4cm]

@@

Currently, a new generation of km-scale (also called `k-scale') resolution
global climate models is in development as the forthcoming phase of climate
modelling. One such model is a 5-km version of the Icosahedral Nonhydrostatic
Weather and Climate Model (ICON), developed jointly by the Deutscher
Wetterdienst (DWD) and Max-Planck-Institute for Meteorology (MPI-M). Because of
the high resolution, most parametrisations, such as those of convection and
clouds, can be avoided.

Previous studies have identified substantial large-scale biases in model clouds
over the Southern Ocean (SO), affecting sea surface temperature and the Earth's
albedo overall. Our aim is to quantify how well the high-resolution ICON model
is simulating clouds in this region, particularly in light of the fact that
subgrid-scale clouds are not parametrised in this model. This region is mostly
dominated by boundary layer clouds generated by shallow convection, and these
are problematic to observe by spaceborne lidar and radars, which are affected
by attenuation by overlapping and thick clouds and ground clutter,
respectively. Therefore, we choose to use a large set of ship-based
observations conducted with ceilometers and lidars on board of the Research
Vessel (RV) \textit{Polarstern} and other voyages. Altogether, we analyse over
2000 days (about 6 years) of data from 31 voyages and 1 sub-antarctic station
covering diverse longitudes of the SO. To achieve a like-for-like comparison
with the model, we use a ground-based lidar simulator called the Automatic
Lidar and Ceilometer Framework (ALCF). We contrast the results with the ECMWF
Reanalysis 5 (ERA5) and the Modern-Era Retrospective analysis for Research and
Applications, Version 2 (MERRA-2).

We show that the model underestimates the total cloud fraction by about 10\%,
with overestimation of cloud below 2 km and underestimation of cloud above 2
km. The reanalyses also underestimate the total cloud fraction by about 20\%.
ERA5 overestimates cloud below 1 km but underestimates near-surface cloud or
fog. In addition to lidar data, we compare radiosonde profiles acquired on the
RV \textit{Polarstern} voyages with ICON. Notably, the model exhibits smaller
natural variability than observations, and its lifting condensation level tends
to be higher. This might explain why cloud occurrence is peaking higher in the
model (at 500 m) than in observations (at the surface).

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/map-fig.png}
\caption{
\textbf{(a)} A map showing the tracks of 31 voyages of RV \emph{Polarstern}, RSV
\emph{Aurora Australis}, RV \emph{Tangaroa}, RV \emph{Nathaniel B. Palmer}, and
HMNZS \emph{Wellington} and one sub-Antarctic station (Macquarie Island)
analysed here. The tracks cover Antarctic sectors south of South America, the
Atlantic Ocean, Africa, Australia, and New Zealand in the years 2010 to 2021.
The dashed line at 40$^\circ$S delineates the Southern Ocean area of our
analysis. A photo of \textbf{(b)} RV \emph{Polarstern} (© Folke
Mehrtens, Alfred-Wegener-Institut), \textbf{(c)} Lufft CHM 15k installed on RV
\emph{Tangaroa} (© Peter Kuma, University of Canterbury), \textbf{(d)} Vaisala
CL51 installation at the Macquarie Island station (© Jeﬀ Aquilina, Bureau of
Meteorology).
% TODO: Ask for photo permission and info about license.
}
\label{fig:map}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{img/lts-dist.pdf}
\caption{
Lower tropospheric stability (LTS) distribution in \textbf{(a)} ERA5 and
\textbf{(b)} MERRA-2 calculated for the 31 voyage tracks and one station from
the highest instantaneous temporal resolution data. Shown is also the chosen
dividing threshold of 12 K for `stable' and `unstable' conditions.
}
\label{fig:lts-dist}
\end{figure}

@@

Missing days in the ceilometer data were HMNZSW16 (7 days): 24--27 November, 10
December, 16--17 December 2016; MARCUS (3 days): 8, 10 November, 10 December;
MCQ16-18 (9 days): 7--8, 29 June, 5, 16 July, 15 August, 17 October 2016, 11
February, 21 March 2017; TAN1502 (1 day): 24 January.

These data sources contained ceilometer (lidar) observations with the Vaisala
CL51 and CL31 operating at 910 nm and the Lufft CHM-15k operating at 1064 nm. A
ceilometer is a low-power vertically pointing lidar with the purpose of
measuring the cloud base, but it also measures the full vertical structure of
clouds as long as the laser signal is not attenuated by thick clouds. A total
of about 1500 days of lidar observations were included in our analysis.

@@

\subsection{Vaisala CL51 and CL31}

The Vaisala CL51 and CL31 are ceilometers operating at a near-infrared
wavelength of 910 nm. The maximum range 15.4 km (CL51), 7.7 km (CL31 with 5 m
vertical resolution) and 7.550 (CL31 with 10 m vertical resolution).
The vertical resolution is 10 m (5 m configurable). The wavelength is affected
by water vapour absorption of about 20\% in the mid-latitudes
\citep{wiegner2015,wiegner2019}.  The instrument data files containing
uncalibrated backscatter were first converted to NetCDF with cl2nc
(\url{https://github.com/peterkuma/cl2nc}) and then processed with the ALCF to
produce a cloud mask and cloud occurrence by height and the total cloud
fraction. CL51 can also be configured to emulate CL31.

For the MARCUS voyage, Vaisala CL31 were made available by the Atmospheric
Radiation Measurement (ARM) processed to a maximum range of 7.545 km, vertical
resolution of 30 m and 252 vertical levels.

@@

The Modern-Era Retrospective analysis for Research and Applications, Version 2
(MERRA-2) is a reanalysis produced by the Global Modeling and Assimilation
Office at the NASA Goddard Space Flight Center \citep{gelaro2017}.  It uses
version 5.12.4 of the Goddard Earth Observing System (GEOS) atmospheric
model \citep{rienecker2008,molod2015}. The reanalysis output analysed here is
available at a spatial resolution of 0.5° in latitude and 0.625° in longitude,
which is about 56 km in the North--South direction and 35 km in the East--West
direction at 60°S. The number of vertical model levels is 72, which we
interpolate to height levels as needed. In our analysis we use the following
products: 1-hourly instantaneous 2D single-level diagnostics (M2I1NXASM) for
2-m temperature and humidity; 3-hourly instantaneous 3D assimilated
meteorological fields (M2I3NVASM) for cloud quantities, pressure and
temperature; 1-hourly average 2D surface flux diagnostics (M2T1NXFLX) for
precipitation and sea ice; and 1-hourly average 2D radiation diagnostics
(M2T1NXRAD) for radiation quantities \citep{merra2}.

@@

ERA5 \citep{era5} is a reanalysis produced by the European Centre for
Medium-Range Weather Forecasts (ECMWF). It is based on a numerical weather
prediction (NWP) model Integrated Forecast System (IFS) version CY41R2.  The
horizontal resolution is 0.25$^\circ$ in latitude and longitude, which is about
28 km in the North--South direction and 14 km in the East--West direction at
60$^\circ$S. The model uses 137 vertical levels. In our analysis we use output
at one hourly instantaeous time intervals, except for radiation quantities,
which are accumulations (from these we calculate daily means).
Vertically-resolved quantities are specified on pressure levels. We interpolate
these quantities to height levels as needed.

@@

\begin{table}
% TODO: Add sampling intervals.
\caption{List of voyages and stations. Start, end and days (UTC; inclusive) indicate the time period when the vessel was south of 40$^\circ$S. Abbreviations: ceilometer (ceil.), Australia (AU), New Zealand (NZ), South America (SA), Atlantic Ocean (AO) and Africa (AF). The number of days is rounded to the nearest integer. CL51/31 indicates CL51 configured to emulate CL31.}
\label{tab:voyages}
\small
\begin{tabular}{llllllr}
Name & Vessel & Ceil. & Region & Start & End & Days\\
\hline
AA15-16  & RSV \emph{Aurora Australis}   & CL51    & AU       & 2015-10-22 & 2016-02-22 & 124\\
HMNZSW16 & HMNZS \emph{Wellington}       & CHM 15k & NZ       & 2016-11-23 & 2016-12-19 & 27 \\
MARCUS   & RSV \emph{Aurora Australis}   & CL31    & AU       & 2017-10-29 & 2018-03-26 & 149\\
MCQ16-18 & /                             & CL51    & AU/NZ    & 2016-04-03 & 2018-03-14 & 710\\
NBP1704  & RV \emph{Nathaniel B. Palmer} & CHM 15k & NZ       & 2017-04-14 & 2017-06-08 & 55 \\
PS77/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2010-12-01 & 2011-02-04 & 65 \\
PS77/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-02-07 & 2011-04-14 & 66 \\
PS79/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2011-12-06 & 2012-01-02 & 27 \\
PS79/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-01-10 & 2012-03-10 & 61 \\
PS79/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-03-14 & 2012-04-08 & 26 \\
PS81/2   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2012-12-02 & 2013-01-18 & 47 \\
PS81/3   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-01-22 & 2013-03-17 & 55 \\
PS81/4   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-03-18 & 2013-04-16 & 30 \\
PS81/5   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-04-20 & 2013-05-23 & 33 \\
PS81/6   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-06-10 & 2013-08-12 & 63 \\
PS81/7   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-08-15 & 2013-10-14 & 60 \\
PS81/8   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-11-12 & 2013-12-14 & 31 \\
PS81/9   & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2013-12-21 & 2014-03-02 & 71 \\
PS89     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2014-12-05 & 2015-01-30 & 56 \\
PS96     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2015-12-08 & 2016-02-14 & 68 \\
PS97     & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-02-15 & 2016-04-06 & 52 \\
PS103    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2016-12-18 & 2017-02-02 & 46 \\
PS104    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2017-02-08 & 2017-03-18 & 39 \\
PS111    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-01-21 & 2018-03-14 & 52 \\
PS112    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-03-18 & 2018-05-05 & 49 \\
PS117    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2018-12-18 & 2019-02-07 & 51 \\
PS118    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2019-02-18 & 2019-04-08 & 50 \\
PS123    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-01-10 & 2021-01-31 & 21 \\
PS124    & RV \emph{Polarstern}          & CL51    & SA/AO/AF & 2021-02-03 & 2021-03-30 & 55 \\
TAN1502  & RV \emph{Tangaroa}            & CL51/31 & NZ       & 2015-01-20 & 2015-03-12 & 51 \\
TAN1702  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2017-03-09 & 2017-03-31 & 23 \\
TAN1802  & RV \emph{Tangaroa}            & CHM 15k & NZ       & 2018-02-07 & 2018-03-20 & 41 \\
\hline
\textbf{Total} &                         &         &          &            &            & \textbf{2421}\\
\end{tabular}
\normalsize
\end{table}
